---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

library(knitr)
     
##Loading and preprocessing the data
####1. Load the data
```{r read_data, echo=TRUE}
unzip("repdata_data_activity.zip")
activity <- read.csv("activity.csv", header=TRUE, sep=",", na.string="NA")
```
####2. Process/transform the data
```{r}
activity_No_NA <- na.omit(activity)
activity_Is_NA <- activity[is.na(activity$steps),]

```
    

##Mean of total number of steps taken per day
####1. Histogram of the total number of steps taken each day
```{r}
NumOfSteps <- aggregate(steps ~ date, activity_No_NA, sum)
plot(NumOfSteps
     , type="h"
     , main ="Total no. of steps taken per day Oct-Nov 2012"
     , xlab="Date"
     , ylab="Number of steps")
 
```

####2.Calculate and report
```{r}
MeanOfSteps <- as.integer(mean(NumOfSteps$steps))
MedianOfSteps <- as.integer(median(NumOfSteps$steps))
```
The mean of total number of steps taken per day is `r MeanOfSteps`.
The median of total number of stpes taken per day is `r MedianOfSteps`.


##The average daily activity pattern
####1. Time series plot
```{r}
NumOfSteps2 <- aggregate(steps ~ interval, activity_No_NA, mean)
plot(NumOfSteps2
     , type="l"
     , main ="Time series plot of Oct-Nov 2012"
     , xlab="5-minute interval"
     , ylab="Average no. of steps taken") 

``` 

####2. Calculate and report 
```{r}
MaxStepsInterval <- NumOfSteps2[NumOfSteps2$steps == max(NumOfSteps2$steps),][[1]]
```
The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is `r MaxStepsInterval`.  




##Imputing missing values
####1. Calculate and report
```{r}
NumOfNA <- sum(is.na(activity$steps))
```
The total number of missing values in the dataset is `r NumOfNA` .

####2. Filling in all of the missing values
```{r}
activity_Is_NA$steps[activity_Is_NA$interval==NumOfSteps2$interval] <- NumOfSteps2$steps
```
####3. Create a new dataset that is equal to the originaldataset but with the missing data filled in
```{r}
activity2 <- rbind(activity_No_NA, activity_Is_NA)
```
####4. Make a histogram, calculate and report
```{r}
NumOfSteps2 <- aggregate(steps ~ date, activity2, sum)
plot(NumOfSteps2
     , type="h"
     , main ="Total no of steps taken per day (missing data was filled with avg interval)"
     , xlab="Date"
     , ylab="Number of steps")

MeanOfSteps2 <- as.integer(mean(NumOfSteps2$steps))
MedianOfSteps2 <- as.integer(median(NumOfSteps2$steps))

```
The mean and median of total number of steps taken per day (missing data was filled with average interval value) are `r MeanOfSteps2` and `r MedianOfSteps2`.  




##Differences in activity patterns between weekdays and weekends
####1. Create weekdays and weekends dataset
```{r}
NumOfSteps3 <- activity2
NumOfSteps3["dayType"] <- ifelse(format(as.Date(NumOfSteps3$date), "%u") >= 6, "Weekend", "Weekday")
weekends <- NumOfSteps3[NumOfSteps3$dayType == "Weekend", c("interval", "steps")]
weekdays <- NumOfSteps3[NumOfSteps3$dayType == "Weekday", c("interval", "steps")]
weekends_Avg <- aggregate(steps ~ interval, weekends, mean)
weekdays_Avg <- aggregate(steps ~ interval, weekdays, mean)
```
####2. Panel plot 
```{r}
par(mfrow = c(2,1), mar = c(4,4,2,1))
plot(weekends_Avg
     , type="l"
     , main ="Weekend"
     , xlab = "5-mintue interval"
     , ylab = "Avg. steps"
     )
plot(weekdays_Avg
     , type="l"
     , main ="Weekday"
     , xlab="5-minute interval"
     , ylab="Avg. Steps")

```
